\section{Conclusion}
\label{sec:conclusion}

We proposed RP-GAAP, a lightweight extension of GAAP that applies rare-pattern weighting to the training loss for graph fraud detection. By computing the rarity of each node's discretized attribute-association pattern and assigning higher sample weights to nodes with infrequent patterns, RP-GAAP improves the model's ability to detect fraud without modifying the GAAP architecture. The method adds three hyperparameters, introduces no learned parameters, and incurs negligible computational overhead.

Experiments on four real-world fraud detection benchmarks show that RP-GAAP improves Recall@K over the original GAAP on YelpChi and T-Finance, with a gain of $+0.97$ points on T-Finance. On Tolokers---the hardest dataset in our evaluation---the combination of rare-pattern weighting and focal loss improves fraud recall by $21.5\%$ relative to the baseline. Elliptic proved largely resistant to loss-level interventions, which we attribute to its temporal graph structure where static feature binning may already produce a sparse pattern space. Hyperparameter sweeps on Tolokers indicate that RP-GAAP is not brittle: all parameter settings produce performance within a narrow band, and no configuration causes catastrophic degradation.

Several limitations motivate future work. First, the current results are from a single random seed; multi-seed evaluation is necessary to establish statistical significance and produce confidence intervals. Second, the rare-pattern weighting mechanism is non-parametric and computed once before training; a learnable or adaptive weighting scheme could potentially capture dataset-specific rarity patterns more effectively. Third, the method's reliance on static feature binning limits its effectiveness on temporal graphs such as Elliptic, suggesting that temporal pattern weighting or sequence-aware discretization could be promising extensions. Fourth, our evaluation is limited to the four datasets for which we reproduced GAAP; extending results to the full set of seven benchmarks from the original GAAP paper would strengthen the empirical case. Finally, combining rare-pattern weighting with architectural improvements such as learned neighbor selection remains an open direction that could yield complementary gains.
